A Bias-Variance Decomposition of Information Transfer in Multi-Task Learning

### 2020-08-30 arxiv/proof_main.tex
wip

	\medskip
	\noindent\textbf{The concentration error between $g(\cW)$ and $\ex{g(\cW)}$.}
	Next, we show that the concentration errors of $g_0, g_1, g_2$ are all lower order terms compared to $\ex{g(\cW)}$.
	For $g_1$, using Lemma \ref{largedeviation} in Appendix \ref{sec_maintools} and the fact that all moments of $\varepsilon_i$ exist, we can obtain the following estimates: for $1\le i , j \le t,$
	$$ |\inner{XB^{\star} \cW^{\top} (\cW\cW^{\top})^{-1}W_j - X\beta_j}{\varepsilon_j}| \le p^\e \sigma\|XB^{\star} \cW^{\top} (\cW\cW^{\top})^{-1}W_j - X\beta_j\|,$$
	and
	\begin{align*}  &|\inner{XB^{\star} \cW^{\top} (\cW\cW^{\top})^{-1}W_j - X\beta_j}{U_X U_X^{\top} \bigbrace{\varepsilon_i W_i^{\top}} (\cW\cW^{\top})^{-1} W_j }| \\
	&\le p^\e \sigma |W_i^{\top} (\cW\cW^{\top})^{-1} W_j|\cdot \|U_X U_X^{\top}( {XB^{\star} \cW^{\top} (\cW\cW^{\top})^{-1}W_j - X\beta_j})\|
	\end{align*}
	with high probability for any constant $\e>0$. Inserting the above estimates into $g_1$, we get that
	\begin{align*} 
	 g_1(\cal W)    &\lesssim p^\e \sigma \sum_{j=1}^t \|  {XB^{\star} \cW^{\top} (\cW\cW^{\top})^{-1}W_j - X\beta_j}\| \\
	&\le  p^\e \sigma \sum_{j=1}^t \|  \Sigma^{1/2}\left(B^{\star} \cW^{\top} (\cW\cW^{\top})^{-1}W_j - \beta_j\right)\| \cdot \|Z\| \\
	&\lesssim p^{1/2+\e} \sigma \sum_{j=1}^t \|  \Sigma_1^{1/2}\left(B^{\star} \cW^{\top} (\cW\cW^{\top})^{-1}W_j - \beta_j\right)\| \\
	& \le p^{1/2+\e} \bignorm{\Sigma^{1/2} B^{\star} (\cW^{\top} (\cW \cW^{\top})^{-1} \cW - \id_{t\times t})}^2 + \sigma^2 p^{1/2+\e} ,
	\end{align*}
	with high probability for any constant $\e>0$. Here in the second step we used $X=Z\Sigma^{1/2}$, in the third step we used the bound \eqref{eq_isometric} to bound the operator norm $\|Z\|$, and in the last step we used the simple AM-GM inequality.

For $g_2$, using Lemma \ref{largedeviation} in Appendix \ref{sec_maintools} and the fact that all moments of $\varepsilon_i$ exist, we can obtain the following estimates: for $1\le i \ne j \le t,$
\begin{align*}
& |\|\e_j\|^2 - n\sigma^2|=| \langle \e_j ,\e_j\rangle - \E\langle \e_j ,\e_j\rangle | \le \sigma^2 p^{1/2+\e},\\
& |\inner{U_X U_X^{\top} \bigbrace{\varepsilon_i W_i^{\top}} (\cW\cW^{\top})^{-1}W_j}{\varepsilon_j}|\le |W_i^{\top} (\cW\cW^{\top})^{-1}W_j| \cdot \sigma^2 p^{\e} \tr \left[(U_X U_X^{\top})^2\right]\le \sigma^2 p^{1/2+\e},\\
 &\left|\left(1-\mathop{\mathbb{E}}_{\varepsilon_j} \right)\left[\inner{U_X U_X^{\top} \bigbrace{\varepsilon_j W_j^{\top}} (\cW\cW^{\top})^{-1}W_j}{\varepsilon_j}\right]\right| \le |W_j^{\top} (\cW\cW^{\top})^{-1}W_j| \cdot \sigma^2 p^{\e} \tr \left[(U_X U_X^{\top})^2\right]\le \sigma^2 p^{1/2+\e},\\
 &\left|\left(1-\mathop{\mathbb{E}}_{\varepsilon_j} \right)\left[\inner{U_X U_X^{\top} \bigbrace{\varepsilon_j W_j^{\top}} (\cW\cW^{\top})^{-1}W_j}{U_X U_X^{\top} \bigbrace{\varepsilon_j W_j^{\top}} (\cW\cW^{\top})^{-1}W_j}\right]\right|\\
 &\le |W_j^{\top} (\cW\cW^{\top})^{-1}W_j|^2 \cdot \sigma^2 p^{\e} \tr \left[(U_X U_X^{\top})^4\right]\le \sigma^2 p^{1/2+\e},
\end{align*}
with high probability for any constant $\e>0$. Inserting the above estimates into $g_1$, we get that
$$\left|g_2(\cal W)-\sigma^2(n\cdot t - p\cdot r)\right|=\left|g_2(\cal W)- \exarg{\varepsilon_1, \dots, \varepsilon_t} {g_2(\cal W)}\right| \lesssim \sigma^2 p^{1/2+\e}$$
with high probability for any constant $\e>0$.

	Finally, for $g_0$ we write it as
	$$g_0(\cal W)= \sum_{j=1}^t\left\|Z v_j \right\|^2,\quad v_j:= \Sigma^{1/2}\left(B^{\star} \cW^{\top} (\cW\cW^{\top})^{-1} W_j -  \beta_j\right).$$
	Note that $Zv_j\in \R^n$ is a random vector with i.i.d. entries of mean zero, variance $\|v_j\|^2$, and finite fourth moment by \eqref{assmAhigh}. Hence by law of large numbers, we have 
	$$\frac{\left\|Z v_j \right\|^2 }{ n \|v_j\|^2}=  1+\oo(1) \quad \text{with probability} \ 1-\oo(1), $$
	which implies that 
	$$g_0(\cal W)=n \sum_{j=1}^t \|v_j\|^2 \cdot (1+\oo(1)) = n \bignorm{\Sigma^{1/2} B^{\star} (\cW^{\top} (\cW \cW^{\top})^{-1} \cW - \id_{t\times t})}^2 \cdot (1+\oo(1))$$
	with probability $1-\oo(1)$.

	Combining all the above estimates, we obtain that for any constant $\e>0$,
	\begin{align*} 
	g(\cal W)&= n \bignorm{\Sigma^{1/2} B^{\star} (\cW^{\top} (\cW \cW^{\top})^{-1} \cW - \id_{t\times t})}^2 \cdot (1+\oo(1)) + \sigma^2 (n\cdot t - p \cdot r)+\OO(\sigma^2 p^{1/2+\e})\\
	&=\ex{g(\cW)} \cdot (1+\oo(1)) 
	\end{align*}
	with probability $1-\oo(1)$.

>>>>

	\medskip
	\noindent\textbf{The concentration error between $g(\cW)$ and $\ex{g(\cW)}$.}
	Next, we show that the concentration errors of $g_0(\cW), g_1(\cW), g_2(\cW)$ are all lower order terms compared to $\ex{g(\cW)}$.

	First, for $g_1(\cW)$, using Lemma \ref{largedeviation} in Appendix \ref{sec_maintools} and the fact that all moments of $\varepsilon_i$ exist, we obtain that for a sufficiently small constant $\e>0$, the following holds with high probability:
	\[ \text{for any } 1\le j \le t, |\inner{XB^{\star} \cW^{\top} (\cW\cW^{\top})^{-1}W_j - X\beta_j}{\varepsilon_j}| \le p^\e \sigma\|XB^{\star} \cW^{\top} (\cW\cW^{\top})^{-1}W_j - X\beta_j\|. \]
	%and
	%\begin{align*}  &|\inner{XB^{\star} \cW^{\top} (\cW\cW^{\top})^{-1}W_j - X\beta_j}{U_X U_X^{\top} \bigbrace{\varepsilon_i W_i^{\top}} (\cW\cW^{\top})^{-1} W_j }| \\
	%&\le p^\e \sigma |W_i^{\top} (\cW\cW^{\top})^{-1} W_j|\cdot \|U_X U_X^{\top}( {XB^{\star} \cW^{\top} (\cW\cW^{\top})^{-1}W_j - X\beta_j})\|
	%\end{align*}
	Applying the above estimates into $g_1(\cW)$, we get that
	\begin{align}
	  \abs{g_1(\cal W)}    &\le p^\e \sigma \sum_{j=1}^t \|  {XB^{\star} \cW^{\top} (\cW\cW^{\top})^{-1}W_j - X\beta_j}\| \nonumber \\
	&\le  p^\e \sigma \sum_{j=1}^t \bignorm{  \Sigma^{1/2}\left(B^{\star} \cW^{\top} (\cW\cW^{\top})^{-1}W_j - \beta_j\right)} \cdot \|Z\| \nonumber\\
	&\lesssim p^{1/2+\e} \sigma \sum_{j=1}^t \bignorm{  \Sigma_1^{1/2}\left(B^{\star} \cW^{\top} (\cW\cW^{\top})^{-1}W_j - \beta_j\right)} \nonumber\\
	& \le p^{1/2+\e} \bignorm{\Sigma^{1/2} B^{\star} (\cW^{\top} (\cW \cW^{\top})^{-1} \cW - \id_{t\times t})}^2 + t\sigma^2 p^{1/2+\e}.  \label{eq_est_g1}
	\end{align}
	In the second step, we use the fact that $X=Z\Sigma^{1/2}$.
	In the third step, we use equation \eqref{eq_isometric} to bound the operator norm $\|Z\|$ by $\OO(\sqrt{p})$.
	In the last step, we use the AM-GM inequality.

	Second, for $g_2(\cW)$, using Lemma \ref{largedeviation} in Appendix \ref{sec_maintools} and the fact that all moments of $\varepsilon_i$ exist, we obtain that for any $1\le i, j \le t$, the following holds with high probability:
	\begin{align*}
		\bigabs{\norm{\varepsilon_j}^2 - \ex{\norm{\varepsilon_j}^2}} &\le \sigma^2 p^{1/2+\e},\\
		%|\inner{U_X U_X^{\top} \bigbrace{\varepsilon_i W_i^{\top}} (\cW\cW^{\top})^{-1}W_j}{\varepsilon_j}|
		\abs{W_i^{\top} (\cW\cW^{\top})^{-1} W_j \cdot \bigbrace{\varepsilon_i^{\top} U_X U_X^{\top} \varepsilon_j}}
		&\le |W_i^{\top} (\cW\cW^{\top})^{-1}W_j| \cdot \sigma^2 p^{\e} \bigtr{(U_X U_X^{\top})^2}^{1/2} \le \sigma^2 p^{1/2+\e}
  % &\left|\left(1-\mathop{\mathbb{E}}_{\varepsilon_j} \right)\left[\inner{U_X U_X^{\top} \bigbrace{\varepsilon_j W_j^{\top}} (\cW\cW^{\top})^{-1}W_j}{\varepsilon_j}\right]\right| \le |W_j^{\top} (\cW\cW^{\top})^{-1}W_j| \cdot \sigma^2 p^{\e} \tr \left[(U_X U_X^{\top})^2\right]\le \sigma^2 p^{1/2+\e},\\
  % &\left|\left(1-\mathop{\mathbb{E}}_{\varepsilon_j} \right)\left[\inner{U_X U_X^{\top} \bigbrace{\varepsilon_j W_j^{\top}} (\cW\cW^{\top})^{-1}W_j}{U_X U_X^{\top} \bigbrace{\varepsilon_j W_j^{\top}} (\cW\cW^{\top})^{-1}W_j}\right]\right|\\
  % &\le |W_j^{\top} (\cW\cW^{\top})^{-1}W_j|^2 \cdot \sigma^2 p^{\e} \tr \left[(U_X U_X^{\top})^4\right]\le \sigma^2 p^{1/2+\e},
	\end{align*}
	Combining the above two inequalities together, we get that with high probability
	\begin{align}
		\left|g_2(\cal W)-\sigma^2(n\cdot t - p\cdot r)\right|=\left|g_2(\cal W)- \exarg{\varepsilon_1, \dots, \varepsilon_t} {g_2(\cal W)}\right| \lesssim \sigma^2 p^{1/2+\e}. \label{eq_est_g2}
	\end{align}

	Finally, for $g_0(\cW)$, denote by $v_j:= \Sigma^{1/2}\left(B^{\star} \cW^{\top} (\cW\cW^{\top})^{-1} W_j -  \beta_j\right)$, for any $j = 1,\dots,t$, we have that $g_0(\cal W)= \sum_{j=1}^t\left\|Z v_j \right\|^2$.
%	$$g_0(\cal W)= \sum_{j=1}^t\left\|Z v_j \right\|^2,\quad v_j:= \Sigma^{1/2}\left(B^{\star} \cW^{\top} (\cW\cW^{\top})^{-1} W_j -  \beta_j\right).$$
	Note that $Zv_j\in \R^n$ is a random vector with i.i.d. entries of mean zero, variance $\|v_j\|^2$, and finite fourth moment by \eqref{assmAhigh}.
	Hence by law of large numbers, we have that with high probability
	${\left\|Z v_j \right\|^2 } = { n \|v_j\|^2} \cdot (1+\oo(1))$,
	which implies that
	\begin{align}
		g_0(\cal W)=n \sum_{j=1}^t \|v_j\|^2 \cdot (1+\oo(1)) = n \bignorm{\Sigma^{1/2} B^{\star} (\cW^{\top} (\cW \cW^{\top})^{-1} \cW - \id_{t\times t})}^2 \cdot (1+\oo(1)). \label{eq_est_g0}
	\end{align}

	Combining equation \eqref{eq_est_g1}, \eqref{eq_est_g2}, and \eqref{eq_est_g0}, we obtain that with high probability
	\begin{align*}
	g(\cal W)&= n \bignorm{\Sigma^{1/2} B^{\star} (\cW^{\top} (\cW \cW^{\top})^{-1} \cW - \id_{t\times t})}^2 \cdot (1+\oo(1)) + \sigma^2 (n\cdot t - p \cdot r)+\OO(\sigma^2 p^{1/2+\e})\\
	&=\ex{g(\cW)} \cdot (1+\oo(1)).
	\end{align*}
